CA of Music Genre Stereotypes

### 2025-06-20 analysis-bb-new.qmd
new bb who dis

```{r import data and select genre stereotype variables}
   #importing data
   dat <- read_dta("C:/Users/Omar Lizardo/Google Drive/MISC DATA SOURCES/SSI-2012/SSI2012.dta")
   #replacing ids with character names
   dat <- bind_cols(tibble(id_char = paste("id", sapply(dat[, 1], as.character), sep = "_")), dat) %>% 
   select(id_char, classicalfemale:hvymtluc, -ends_with("none")) %>%  #selecting relevant columns
   na.omit() #removing missing values
```

```{r Creating individual matrices}
   row.breiger <- function(x) {
      x %*% t(x)
      }
   col.breiger <- function(x) {
      t(x) %*% x
      }
   bb <- function(x) {
      x <- suppressMessages(disparity(x, alpha = 0.25, narrative = FALSE))
      return(x)
      }
   den <- rowSums(dat[,2:301])/300
   bad.den <- which(den > .9 | den < 0.1)
   dat2 <- dat[-c(bad.den), 2:301]
   ind.list <- split(dat2, seq(nrow(dat2))) # splitting each row of data frame into list objects
   ind.mat <- lapply(ind.list, function(x) {matrix(as.numeric(x), 20, 15, byrow = TRUE)}) # creating individual two mode network of genre by labels
      r <- c("Classical", "Opera", "Jazz", "Broadway", "Easy", "Big_Band", "Classic_Rock", "Country", "Bluegrass", "Folk", "Gospel", "Latin", "Rap", "Blues", "Reggae", "Pop", "Contemp_Rock", "Alt_Rock", "Dance", "Metal")
       c <- c("Female", "Male", "White", "Black", "Hispanic", "Asian", "College", "No_College", "Young", "Mid_Age", "Old", "Lower_Class", "Working_Class", "Middle_Class", "Upper_Class")
   ind.mat <- lapply(ind.mat, function(x) {rownames(x) <- r; x}) # naming individual rows
   ind.mat <- lapply(ind.mat, function(x) {colnames(x) <- c; x}) # naming individual columns
   names(ind.mat) <- dat$id_char[-c(bad.den)]
   ind.row <- lapply(ind.mat, row.breiger) # individual genre projection
   ind.col <- lapply(ind.mat, col.breiger) # individual label projection
   row.bb <- lapply(ind.row, bb)
   col.bb <- lapply(ind.col, bb)
   row.stack <- matrix(unlist(row.bb), ncol = 20, byrow = TRUE) # genre stacked matrix
   col.stack <- matrix(unlist(col.bb), ncol = 15, byrow = TRUE) # label stacked matrix
   colnames(row.stack) <- r # naming genre stacked matrix columns
   colnames(col.stack) <- c # naming label stacked matrix columns
   rownames(row.stack) <- rep(r, nrow(dat2)) # naming genre stacked matrix rows
   rownames(col.stack) <- rep(c, nrow(dat2)) # naming label stacked matrix rows
   zero.rows.r <- which(rowSums(row.stack) == 0) # which genre stacked matrix rows sum to zero?
   zero.rows.c <- which(rowSums(col.stack) == 0) # which label stacked matrix rows sum to zero?
   names(ind.row) <- dat$id_char[-c(bad.den)]
   names(ind.col) <- dat$id_char[-c(bad.den)]
   names(row.bb) <- dat$id_char[-c(bad.den)]
   names(col.bb) <- dat$id_char[-c(bad.den)]
```

```{r Stacked CA}
   #genres
   row.stack.d <- data.frame(row.stack)
   row.stack.d$genre <- rep(r, nrow(dat2))
   row.stack.d$id <- rep(dat$id_char[-bad.den], each = 20)
   CA.r <- CA(row.stack.d, graph = FALSE, row.sup = zero.rows.r, quali.sup = c(21, 22))
   #labels
   col.stack.d <- data.frame(col.stack)
   col.stack.d$label <- rep(c, nrow(dat2))
   col.stack.d$id <- rep(dat$id_char[-bad.den], each = 15)
   CA.c <- CA(col.stack.d, graph = FALSE, row.sup = zero.rows.c, quali.sup = c(16, 17))
```

@@

```{r Plot Data}
     make.plot.dat <- function(x) { 
         w <-  data.frame(x) %>% 
         mutate(col.obj = gsub("\\..*","",rownames(x))) 
         rownames(w) <- 1:nrow(x)
         return(w)
         }
      col.gen.dat <-  data.frame(CA.r$col$coord) %>% 
         mutate(col.obj = rownames(CA.r$col$coord)) #genre column scores
      col.lab.dat <-  data.frame(CA.c$col$coord) %>% 
         mutate(col.obj = rownames(CA.c$col$coord)) #label column scores
      sup.gen.dat <- CA.r$quali.sup$coord[1:20, ] 
      sup.lab.dat <- CA.c$quali.sup$coord[1:15, ] 
      sup.gen.dat <- data.frame(sup.gen.dat, 
                                col.obj = gsub("genre.", "", rownames(sup.gen.dat)))
      rownames(sup.gen.dat) <- sup.gen.dat$col.obj
      sup.lab.dat <- data.frame(sup.lab.dat, 
                                col.obj = gsub("label.", "", rownames(sup.lab.dat)))
      rownames(sup.lab.dat) <- sup.lab.dat$col.obj
      row.gen.dat <- make.plot.dat(CA.r$row$coord) #genre row scores
      row.lab.dat <- make.plot.dat(CA.c$row$coord) #label row scores
   ind.gen.sup <- CA.r$quali.sup$coord %>% 
      data.frame() %>% 
      slice(21:nrow(CA.r$quali.sup$coord)) 
   ind.gen.sup$id_char <- str_sub(rownames(ind.gen.sup), 4, -1)
   ind.lab.sup <- CA.c$quali.sup$coord %>% 
      data.frame() %>% 
      slice(21:nrow(CA.c$quali.sup$coord)) 
   ind.lab.sup$id_char <- str_sub(rownames(ind.lab.sup), 4, -1)
```

>>>>

```{r import data and select genre stereotype variables}
   #importing data
   dat <- read_dta("C:/Users/Omar Lizardo/Google Drive/MISC DATA SOURCES/SSI-2012/SSI2012.dta")
   #replacing ids with character names
   dat <- bind_cols(tibble(id_char = paste("id", sapply(dat[, 1], as.character), sep = "_")), dat) %>% 
   select(id_char, classicalfemale:hvymtluc, -ends_with("none")) %>%  #selecting relevant columns
   na.omit() #removing missing values
      r <- c("Classical", "Opera", "Jazz", "Broadway", "Easy", "Big_Band", "Classic_Rock", "Country", "Bluegrass", "Folk", "Gospel", "Latin", "Rap", "Blues", "Reggae", "Pop", "Contemp_Rock", "Alt_Rock", "Dance", "Metal")
       c <- c("Female", "Male", "White", "Black", "Hispanic", "Asian", "College", "No_College", "Young", "Mid_Age", "Old", "Lower_Class", "Working_Class", "Middle_Class", "Upper_Class")
```

```{r Summary Table of Averaged Perceptions}
   tab <- colSums(dat[,2:301])/nrow(dat)
   tab <- matrix(tab, nrow = 20,  byrow = TRUE)
   colnames(tab) <- c
   rownames(tab) <- r
   png(file = here("Plots", "main.png"), width=1000, height=700)
   corrplot(tab, is.corr = FALSE, method = "color", 
      tl.col = "black", tl.cex = 1, tl.srt = 45,
      addCoef.col = "black", cl.pos = "n", cl.lim = c(0, max(tab)),
      number.cex = 1, number.digits = 2, addgrid.col = "darkgray",
      col=colorRampPalette(c("lightblue","orange","red"))(200))
   dev.off()
```

```{r Creating individual matrices}
   make.mat <- function(x) {
      x <- matrix(as.numeric(x), 20, 15, byrow = TRUE)
      return(x)
      }
   bb.sdsm <- function(x) {
     tryCatch({
         w <- sdsm(x, alpha = 0.25, narrative = FALSE)
         return(w)
            }, error = function(e) {return(NULL)})
      }
   den <- rowSums(dat[,2:301])/300
   bad.den <- which(den > .9 | den < 0.1)
   dat2 <- dat[-c(bad.den), 2:301]
   ind.list <- split(dat2, seq(nrow(dat2))) # splitting each row of data frame into list objects
   ind.mat <- lapply(ind.list, make.mat) # creating individual two mode network of genre by labels
   ind.mat <- lapply(ind.mat, function(x) {rownames(x) <- r; x}) # naming individual rows
   ind.mat <- lapply(ind.mat, function(x) {colnames(x) <- c; x}) # naming individual columns
   names(ind.mat) <- dat$id_char[-c(bad.den)]
   ind.mat.t <- lapply(ind.mat, function(x) {t(x)})
   row.bb <- lapply(ind.mat, bb.sdsm)
   col.bb <- lapply(ind.mat.t, bb.sdsm)
   null.rows <- which(as.numeric(unlist(lapply(row.bb, is.null))) == 1)
   null.cols <- which(as.numeric(unlist(lapply(col.bb, is.null))) == 1)
   row.bb <- row.bb[-null.rows]
   col.bb <- col.bb[-null.cols]
   row.stack <- matrix(unlist(row.bb), ncol = 20, byrow = TRUE) # genre stacked matrix
   col.stack <- matrix(unlist(col.bb), ncol = 15, byrow = TRUE) # label stacked matrix
   colnames(row.stack) <- r # naming genre stacked matrix columns
   colnames(col.stack) <- c # naming label stacked matrix columns
   m <- nrow(row.stack)/20
   rownames(row.stack) <- rep(r, m) # naming genre stacked matrix rows
   rownames(col.stack) <- rep(c, m) # naming label stacked matrix rows
   zero.rows.r <- which(rowSums(row.stack) == 0) # which genre stacked matrix rows sum to zero?
   zero.rows.c <- which(rowSums(col.stack) == 0) # which label stacked matrix rows sum to zero?
```

```{r Stacked CA}
   #genres
   row.stack.d <- data.frame(row.stack)
   row.stack.d$genre <- rep(r, m)
   row.stack.d$id <- rep(names(row.bb), each = 20)
   CA.r <- CA(row.stack.d, graph = FALSE, row.sup = zero.rows.r, quali.sup = c(21, 22))
   #labels
   col.stack.d <- data.frame(col.stack)
   col.stack.d$label <- rep(c, m)
   col.stack.d$id <- rep(names(col.bb), each = 15)
   CA.c <- CA(col.stack.d, graph = FALSE, row.sup = zero.rows.c, quali.sup = c(16, 17))
```

@@

```{r Plot Data}
     make.plot.dat <- function(x) { 
         w <-  data.frame(x) %>% 
         mutate(col.obj = gsub("\\..*","",rownames(x))) 
         rownames(w) <- 1:nrow(x)
         return(w)
         }
      #genre column scores data frame
      col.gen.dat <-  data.frame(CA.r$col$coord) %>% 
         mutate(col.obj = rownames(CA.r$col$coord)) 
      #label column scores data frame
      col.lab.dat <-  data.frame(CA.c$col$coord) %>% 
         mutate(col.obj = rownames(CA.c$col$coord)) 
      #supplementary scores data frame for genres
      sup.gen.dat <- CA.r$quali.sup$coord[1:20, ] 
      sup.gen.dat <- data.frame(sup.gen.dat, 
                                col.obj = gsub("genre.", "", rownames(sup.gen.dat)))
      rownames(sup.gen.dat) <- 1:nrow(sup.gen.dat)
      #supplementary scores data frame for labels
      sup.lab.dat <- CA.c$quali.sup$coord[1:15, ] 
      sup.lab.dat <- data.frame(sup.lab.dat, 
                                col.obj = gsub("label.", "", rownames(sup.lab.dat)))
      rownames(sup.lab.dat) <- 1:nrow(sup.lab.dat)
      #genre row scores
      row.gen.dat <- make.plot.dat(CA.r$row$coord) 
      #label row scores
      row.lab.dat <- make.plot.dat(CA.c$row$coord) 
      #supplementary individual scores for genres
      ind.gen.sup <- CA.r$quali.sup$coord %>% 
         data.frame() %>% 
         slice(21:nrow(CA.r$quali.sup$coord)) %>% 
         mutate(id_char = str_sub(rownames(CA.r$quali.sup$coord[21:nrow(CA.r$quali.sup$coord), ]), 4, -1))
      row.names(ind.gen.sup) <- 1:nrow(ind.gen.sup)
      #supplementary individual scores for labels
      ind.lab.sup <- CA.c$quali.sup$coord %>% 
         data.frame() %>% 
         slice(16:nrow(CA.c$quali.sup$coord)) %>% 
         mutate(id_char = str_sub(rownames(CA.c$quali.sup$coord[16:nrow(CA.c$quali.sup$coord), ]), 4, -1))
      row.names(ind.lab.sup) <- 1:nrow(ind.lab.sup)
```
